# Replaying disequilibrium populations

The paper shows fine-grained replays for three replicates that were experiencing adaptive momentum (one that failed to cross, one that crossed one valley, and one that crossed two valleys).

Prior to these fine-grained replays, we replayed 24 populations at with less precision. 
We replayed every fourth generation with 1,000 replicates per replay (compared to 10,000 replicates for every generation in the fine-grained replays). 
Here we show all 24 coarse-grained replays.

The functions to generate the plots are not shown here. 
They can be viewed in `/experiments/2024_03_18_01__replays/analysis`

For each plot, the top subplot shows the potential to cross the first valley (orange line) and the second valley (yellow line).
The background of the top subplot shows the expectation based on the position of the leading edge, as in the paper.
The bottom subplot shows the state of the population at each generation of the original replicate. 

## Dependencies

```{r, message=FALSE}
# External
library(ggplot2)
library(dplyr)
library(cowplot)

base_repo_dir = '../..'
exp_dir = paste0(base_repo_dir, '/experiments/2024_03_18_01__replays/')

# Internal
source(paste0(base_repo_dir, '/global_shared_files/global_analysis_variables.R'))
```

```{r, message=FALSE, include=FALSE}
adjust_benchmarking = function(df_benchmarking, df_snapshot){
  df_adjusted_benchmarking = data.frame(data = matrix(nrow = 0, ncol = 4))
  colnames(df_adjusted_benchmarking) = c('update', 'leading_edge_val', 'leading_edge_index', 'frac_crossed')
  for(leading_edge_val in unique(df_benchmarking$leading_edge_val)){
    prev_leading_edge_index = -1
    for(update in sort(unique(df_snapshot$update))){
      leading_edge_index = df_snapshot[df_snapshot$update == update,]$leading_edge_index
      if(leading_edge_index != prev_leading_edge_index){
        frac_crossed = df_benchmarking[
          df_benchmarking$leading_edge_index <= leading_edge_index & 
            df_benchmarking$leading_edge_index > leading_edge_index - 8 & 
            df_benchmarking$leading_edge_val == leading_edge_val,]$crossed_frac
        df_adjusted_benchmarking[nrow(df_adjusted_benchmarking) + 1,] = c(
          update, 
          leading_edge_val, 
          leading_edge_index,
          frac_crossed
        )
        prev_leading_edge_index = leading_edge_index
      }
    }
  }
  return(df_adjusted_benchmarking) 
}

adjust_benchmarking_for_second_cross = function(df_benchmarking, df_snapshot){
  df_adjusted_benchmarking = data.frame(data = matrix(nrow = 0, ncol = 9))
  colnames(df_adjusted_benchmarking) = c('update', 
                                         'left_leading_edge_val', 'left_leading_edge_index', 'left_frac_crossed', 'true_left_val',
                                         'right_leading_edge_val', 'right_leading_edge_index', 'right_frac_crossed', 'true_right_val')
  for(update in sort(unique(df_snapshot$update))){
    snapshot_row = df_snapshot[df_snapshot$update == update,]
    if(snapshot_row$second_leading_edge_left_index == 'None' | update >= 768){
      df_adjusted_benchmarking[nrow(df_adjusted_benchmarking) + 1,] = c( update, NA, NA, NA, NA, NA, NA, NA, NA)
      next
    }
    updates_remaining = 768 - update
    # Calculate cross probability on the left edge (remember it's moving left)
    left_index = as.numeric(snapshot_row$second_leading_edge_left_index)
    # Invert the direction
    left_index = 512 - left_index
    left_index = max(left_index, (512 - updates_remaining * 0.66))
    closest_left_index = left_index - (left_index %% 8)
    true_left_val = as.numeric(snapshot_row$second_leading_edge_left_val)
    left_frac_crossed = 0
    # Calculate cross probability on the right edge
    right_index = as.numeric(snapshot_row$second_leading_edge_right_index)
    right_index = max(right_index, 512 - updates_remaining)
    closest_right_index = right_index - (right_index %% 8)
    true_right_val = as.numeric(snapshot_row$second_leading_edge_right_val)
    for(val in 12:17){
      left_frac_crossed = df_benchmarking[df_benchmarking$leading_edge_index == closest_left_index & df_benchmarking$leading_edge_val == val,]$crossed_frac
      right_frac_crossed = df_benchmarking[df_benchmarking$leading_edge_index == closest_right_index & df_benchmarking$leading_edge_val == val,]$crossed_frac
      df_adjusted_benchmarking[nrow(df_adjusted_benchmarking) + 1,] = c(
        update, 
        val, left_index, left_frac_crossed, true_left_val,
        val, right_index, right_frac_crossed, true_right_val
      )
    }
  }
  #df_adjusted_benchmarking$full_estimate = 1 - ((1 - df_adjusted_benchmarking$left_frac_crossed) * (1 - df_adjusted_benchmarking$right_frac_crossed))
  return(df_adjusted_benchmarking) 

}

pivot_snapshot_data_longer = function(df_snapshot){
  df_snapshot_longer = tidyr::pivot_longer(df_snapshot, paste0('count_', c(6:18, 'under', 'over')), names_to = 'category')
  df_snapshot_longer$category_factor = factor(df_snapshot_longer$category, levels = paste0('count_', c('under', 6:18, 'over')))
  return(df_snapshot_longer)
}

load_and_reformat_full_snapshot_data = function(rep_id){
  df_full_snapshot_data = read.csv(paste0(exp_dir, '/data/reps/', rep_id, '/full_snapshot_data.csv'))
  df_full_snapshot_longer = tidyr::pivot_longer(df_full_snapshot_data, paste0('idx_', 0:511), 'index', names_prefix = 'idx_')
  df_full_snapshot_longer$index = as.numeric(df_full_snapshot_longer$index)
  df_full_snapshot_longer$value_str = as.character(df_full_snapshot_longer$value)
  df_full_snapshot_longer[df_full_snapshot_longer$value < 6,]$value_str = 'under'
  df_full_snapshot_longer[df_full_snapshot_longer$value > 24,]$value_str = 'over'
  df_full_snapshot_longer$value_factor = factor(df_full_snapshot_longer$value_str, levels = paste0(c('under', 6:24, 'over')))
  return(df_full_snapshot_longer)
}

plot_replay_with_adjusted_benchmark = function(rep_id){

  processed_data_dir = paste0(exp_dir, '/data/processed')
  # Load processed data
  df_rep_summary = read.csv(paste0(processed_data_dir, '/processed_summary_rep_', rep_id, '.csv'))  
  df_second_cross_summary = read.csv(paste0(processed_data_dir, '/processed_second_cross_summary_rep_', rep_id, '.csv'))  
  # Load and prep benchmarking data (as in script 01)
  df_benchmarking = read.csv(paste0(base_repo_dir, '/experiments/2024_03_14_01__benchmarking_10k/data/processed_summary.csv'))
  
  # Load snapshot data and also create a longer form "tidy" version
  df_snapshot = read.csv(paste0(exp_dir, '/data/reps/', rep_id, '/snapshot_data.csv'))
  df_snapshot_longer = pivot_snapshot_data_longer(df_snapshot)
  
  # Leverage snapshot data to adjust benchmarking data
  df_adjusted_benchmarking = adjust_benchmarking(df_benchmarking, df_snapshot) 
  # Pivot the benchmarking data to be wider so we can plot it as an area 
  df_benchmarking_wide = tidyr::pivot_wider(df_adjusted_benchmarking, names_prefix = 'frac_crossed_', names_from = c(leading_edge_val), values_from = c(frac_crossed))
  
  # Create a combined potentiation + muller plot
  df_full_snapshot = load_and_reformat_full_snapshot_data(rep_id)
  first_cross_ud = min(df_full_snapshot[df_full_snapshot$value >= 18,]$update)
  
  line_size = 0.75
  ggp_base_area = ggplot(df_benchmarking_wide, aes(x = update)) +
    geom_ribbon(aes(ymin = 0, ymax = frac_crossed_12 * 100, fill = '12'), alpha = 1) + 
    geom_ribbon(aes(ymin = frac_crossed_12 * 100, ymax = frac_crossed_13 * 100, fill = '13'), alpha = 1) + 
    geom_ribbon(aes(ymin = frac_crossed_13 * 100, ymax = frac_crossed_14 * 100, fill = '14'), alpha = 1) + 
    geom_ribbon(aes(ymin = frac_crossed_14 * 100, ymax = frac_crossed_15 * 100, fill = '15'), alpha = 1) + 
    geom_ribbon(aes(ymin = frac_crossed_15 * 100, ymax = frac_crossed_16 * 100, fill = '16'), alpha = 1) + 
    geom_ribbon(aes(ymin = frac_crossed_16 * 100, ymax = frac_crossed_17 * 100, fill = '17'), alpha = 1) + 
    geom_vline(aes(xintercept=first_cross_ud), linetype = 'dashed', alpha = 0.5) + 
    geom_line(data = df_adjusted_benchmarking, aes(x = update, y = frac_crossed * 100, color = as.factor(leading_edge_val))) + 
    geom_line(data = df_rep_summary, aes(x = replay_gen, y = frac_crossed * 100), color = color_map_value['18'], size = line_size) + 
    geom_line(data = df_second_cross_summary, aes(x = replay_gen, y = frac_crossed * 100), color = color_map_value['24'], size = line_size) + 
    scale_y_continuous(limits = c(0,100), expand = c(0,0)) + 
    scale_x_continuous(limits = c(0, 768), expand = c(0,0)) + 
    scale_fill_manual(values = color_map_value) +
    scale_color_manual(values = color_map_value) + 
    xlab('Time (generations)') + 
    ylab('Potential to cross') +
    theme(axis.title = element_text(size = 16)) + 
    theme(axis.text = element_text(size = 14)) + 
    theme(legend.position = 'none') + 
    theme(panel.background = element_rect(fill = '#bbbbbb')) +
    theme(panel.grid.major = element_line(color = '#aaaaaa')) +
    theme(panel.grid.minor = element_line(color = '#aaaaaa'))
  
  ggp_muller = ggplot(df_full_snapshot, aes(x = update, y = index, fill = value_factor)) + 
    geom_raster() + 
    geom_vline(aes(xintercept=first_cross_ud), linetype = 'dashed', alpha = 0.5) + 
    xlab('Time (generations)') + 
    ylab('Index in population') + 
    labs(fill = 'Value') + 
    scale_fill_manual(values = color_map_value) + 
    scale_x_continuous(limits = c(0, 768), expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0)) + 
    theme(axis.title = element_text(size = 16)) + 
    theme(axis.text = element_text(size = 14)) + 
    theme(legend.position = 'none')

  combined_plot = cowplot::plot_grid(ggp_base_area, ggp_muller, nrow = 2, ncol = 1)
  show(combined_plot)
}

```

## Populations that did not cross
 
### Seed 134
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('134')
```
 
### Seed 158
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('158')
```
 
### Seed 164
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('164')
```

### Seed 175
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('175')
```
 
### Seed 252
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('252')
```
 
### Seed 339

Note that this is the seed in Figure 6 of the paper.

```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('339')
```
 
### Seed 365
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('365')
```
 
### Seed 394
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('394')
```
 
### Seed 446
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('446')
```
 
### Seed 450
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('450')
```


## Populations that crossed once

### Seed 011
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('011')
```

### Seed 050
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('050')
```

### Seed 075
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('075')
```

### Seed 083
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('083')
```

### Seed 105
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('105')
```

### Seed 282
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('282')
```

### Seed 343
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('343')
```

### Seed 400

Note that this is the seed in Figure 4 of the paper.

```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('400')
```

### Seed 408
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('408')
```

### Seed 415
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('415')
```


## Populations that crossed twice
 
### Seed 093
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('093')
```

### Seed 124
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('124')
```

### Seed 138
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('138')
```

### Seed 263

Note that this is the seed in Figure 5 of the paper.

```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('263')
```


