# Replaying equilibrium populations

The paper shows fine-grained replays for one replicate that was NOT experiencing adaptive momentum at the start.
The replicate shown in the paper crossed two valleys, and was the only one to do so. 
We also replayed 10 randomlyselected replicates that crossed one valley and 10 randomly selected replicates that failed to cross any valleys. 
We show those here. 

The functions to generate the plots are not shown here. 
They can be viewed in `/experiments/2024_03_22_01__no_am_replays/analysis

For each plot, the top subplot shows the potential to cross the first valley (red line) and the second valley (orange line).
The background of the top subplot shows the expectation based on the position of the leading edge, as in the paper.
The bottom subplot shows the state of the population at each generation of the original replicate. 

## Dependencies

```{r, message=FALSE}
# External
library(ggplot2)
library(dplyr)
library(cowplot)

base_repo_dir = '../..'
exp_dir = paste0(base_repo_dir, '/experiments/2024_03_22_01__no_am_replays/')

# Internal
source(paste0(base_repo_dir, '/global_shared_files/global_analysis_variables.R'))
```

```{r, message=FALSE, include=FALSE}
adjust_benchmarking = function(df_benchmarking, df_snapshot){
  df_adjusted_benchmarking = data.frame(data = matrix(nrow = 0, ncol = 4))
  colnames(df_adjusted_benchmarking) = c('update', 'leading_edge_val', 'leading_edge_index', 'frac_crossed')
  for(leading_edge_val in unique(df_benchmarking$leading_edge_val)){
    prev_leading_edge_index = -1
    for(update in sort(unique(df_snapshot$update))){
      leading_edge_index = df_snapshot[df_snapshot$update == update,]$leading_edge_index
      if(leading_edge_index != prev_leading_edge_index){
        frac_crossed = df_benchmarking[
          df_benchmarking$leading_edge_index <= leading_edge_index & 
            df_benchmarking$leading_edge_index > leading_edge_index - 8 & 
            df_benchmarking$leading_edge_val == leading_edge_val,]$crossed_frac
        df_adjusted_benchmarking[nrow(df_adjusted_benchmarking) + 1,] = c(
          update, 
          leading_edge_val, 
          leading_edge_index,
          frac_crossed
        )
        prev_leading_edge_index = leading_edge_index
      }
    }
  }
  return(df_adjusted_benchmarking) 
}

adjust_benchmarking_for_second_cross = function(df_benchmarking, df_snapshot){
  df_adjusted_benchmarking = data.frame(data = matrix(nrow = 0, ncol = 9))
  colnames(df_adjusted_benchmarking) = c('update', 
                                         'left_leading_edge_val', 'left_leading_edge_index', 'left_frac_crossed', 'true_left_val',
                                         'right_leading_edge_val', 'right_leading_edge_index', 'right_frac_crossed', 'true_right_val')
  for(update in sort(unique(df_snapshot$update))){
    snapshot_row = df_snapshot[df_snapshot$update == update,]
    if(snapshot_row$second_leading_edge_left_index == 'None' | update >= 768){
      df_adjusted_benchmarking[nrow(df_adjusted_benchmarking) + 1,] = c( update, NA, NA, NA, NA, NA, NA, NA, NA)
      next
    }
    updates_remaining = 768 - update
    # Calculate cross probability on the left edge (remember it's moving left)
    left_index = as.numeric(snapshot_row$second_leading_edge_left_index)
    # Invert the direction
    left_index = 512 - left_index
    left_index = max(left_index, (512 - updates_remaining * 0.66))
    closest_left_index = left_index - (left_index %% 8)
    true_left_val = as.numeric(snapshot_row$second_leading_edge_left_val)
    left_frac_crossed = 0
    # Calculate cross probability on the right edge
    right_index = as.numeric(snapshot_row$second_leading_edge_right_index)
    right_index = max(right_index, 512 - updates_remaining)
    closest_right_index = right_index - (right_index %% 8)
    true_right_val = as.numeric(snapshot_row$second_leading_edge_right_val)
    for(val in 12:17){
      left_frac_crossed = df_benchmarking[df_benchmarking$leading_edge_index == closest_left_index & df_benchmarking$leading_edge_val == val,]$crossed_frac
      right_frac_crossed = df_benchmarking[df_benchmarking$leading_edge_index == closest_right_index & df_benchmarking$leading_edge_val == val,]$crossed_frac
      df_adjusted_benchmarking[nrow(df_adjusted_benchmarking) + 1,] = c(
        update, 
        val, left_index, left_frac_crossed, true_left_val,
        val, right_index, right_frac_crossed, true_right_val
      )
    }
  }
  #df_adjusted_benchmarking$full_estimate = 1 - ((1 - df_adjusted_benchmarking$left_frac_crossed) * (1 - df_adjusted_benchmarking$right_frac_crossed))
  return(df_adjusted_benchmarking) 

}

pivot_snapshot_data_longer = function(df_snapshot){
  df_snapshot_longer = tidyr::pivot_longer(df_snapshot, paste0('count_', c(6:18, 'under', 'over')), names_to = 'category')
  df_snapshot_longer$category_factor = factor(df_snapshot_longer$category, levels = paste0('count_', c('under', 6:18, 'over')))
  return(df_snapshot_longer)
}

load_and_reformat_full_snapshot_data = function(rep_id){
  df_full_snapshot_data = read.csv(paste0(exp_dir, '/data/reps/', rep_id, '/full_snapshot_data.csv'))
  df_full_snapshot_longer = tidyr::pivot_longer(df_full_snapshot_data, paste0('idx_', 0:511), 'index', names_prefix = 'idx_')
  df_full_snapshot_longer$index = as.numeric(df_full_snapshot_longer$index)
  df_full_snapshot_longer$value_str = as.character(df_full_snapshot_longer$value)
  df_full_snapshot_longer[df_full_snapshot_longer$value < 6,]$value_str = 'under'
  df_full_snapshot_longer[df_full_snapshot_longer$value > 24,]$value_str = 'over'
  df_full_snapshot_longer$value_factor = factor(df_full_snapshot_longer$value_str, levels = paste0(c('under', 6:24, 'over')))
  df_full_snapshot_longer$value_remap = df_full_snapshot_longer$value - 6
  df_full_snapshot_longer$value_remap_str = as.character(df_full_snapshot_longer$value_remap)
  df_full_snapshot_longer[df_full_snapshot_longer$value_remap < 6,]$value_remap_str = 'under'
  df_full_snapshot_longer[df_full_snapshot_longer$value_remap > 24,]$value_remap_str = 'over'
  df_full_snapshot_longer$value_remap_factor = factor(df_full_snapshot_longer$value_remap_str, levels = paste0(c('under', 6:24, 'over')))
  return(df_full_snapshot_longer)
}

plot_replay_with_adjusted_benchmark = function(rep_id){

  processed_data_dir = paste0(exp_dir, '/data/processed')
  # Load processed data
  df_rep_summary = read.csv(paste0(processed_data_dir, '/processed_summary_rep_', rep_id, '.csv'))  
  df_second_cross_summary = read.csv(paste0(processed_data_dir, '/processed_second_cross_summary_rep_', rep_id, '.csv'))  
  # Load and prep benchmarking data (as in script 01)
  df_benchmarking = read.csv(paste0(base_repo_dir, '/experiments/2024_03_14_01__benchmarking_10k/data/processed_summary.csv'))
  
  # Load snapshot data and also create a longer form "tidy" version
  df_snapshot = read.csv(paste0(exp_dir, '/data/reps/', rep_id, '/snapshot_data.csv'))
  df_snapshot_longer = pivot_snapshot_data_longer(df_snapshot)
  
  # Leverage snapshot data to adjust benchmarking data
  df_adjusted_benchmarking = adjust_benchmarking(df_benchmarking, df_snapshot) 
  # Pivot the benchmarking data to be wider so we can plot it as an area 
  df_benchmarking_wide = tidyr::pivot_wider(df_adjusted_benchmarking, names_prefix = 'frac_crossed_', names_from = c(leading_edge_val), values_from = c(frac_crossed))
  
  # Create a combined potentiation + muller plot
  df_full_snapshot = load_and_reformat_full_snapshot_data(rep_id)
  first_cross_ud = min(df_full_snapshot[df_full_snapshot$value >= 12,]$update)
  
  line_size = 0.75
  ggp_base_area = ggplot(df_benchmarking_wide, aes(x = update)) +
    geom_vline(aes(xintercept=first_cross_ud), linetype = 'dashed', alpha = 0.5) + 
    geom_line(data = df_adjusted_benchmarking, aes(x = update, y = frac_crossed * 100, color = as.factor(leading_edge_val))) + 
    geom_line(data = df_rep_summary, aes(x = replay_gen, y = frac_crossed * 100), color = color_map_value['12'], size = line_size) + 
    geom_line(data = df_second_cross_summary, aes(x = replay_gen, y = frac_crossed * 100), color = color_map_value['18'], size = line_size) + 
    scale_y_continuous(limits = c(0,100), expand = c(0,0)) + 
    scale_x_continuous(limits = c(0, 768), expand = c(0,0)) + 
    scale_fill_manual(values = color_map_value) +
    scale_color_manual(values = color_map_value) + 
    xlab('Time (generations)') + 
    ylab('Potential to cross') +
    theme(axis.title = element_text(size = 16)) + 
    theme(axis.text = element_text(size = 14)) + 
    theme(legend.position = 'none') 
    #theme(panel.background = element_rect(fill = '#bbbbbb')) +
    #theme(panel.grid.major = element_line(color = '#aaaaaa')) +
    #theme(panel.grid.minor = element_line(color = '#aaaaaa'))
  
  ggp_muller = ggplot(df_full_snapshot, aes(x = update, y = index, fill = value_remap_factor)) + 
    geom_raster() + 
    geom_vline(aes(xintercept=first_cross_ud), linetype = 'dashed', alpha = 0.5) + 
    xlab('Time (generations)') + 
    ylab('Index in population') + 
    labs(fill = 'Value') + 
    scale_fill_manual(values = color_map_value) + 
    scale_x_continuous(limits = c(0, 768), expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0)) + 
    theme(axis.title = element_text(size = 16)) + 
    theme(axis.text = element_text(size = 14)) + 
    theme(legend.position = 'none')

  combined_plot = cowplot::plot_grid(ggp_base_area, ggp_muller, nrow = 2, ncol = 1)
  show(combined_plot)
}

```

## Populations that did not cross
 
### Seed 01164
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('01164')
```

### Seed 01435
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('01435')
```

### Seed 01572
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('01572')
```

### Seed 02581
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('02581')
```

### Seed 02711
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('02711')
```

### Seed 02961
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('02961')
```

### Seed 04390
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('04390')
```

### Seed 06116
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('06116')
```

### Seed 06583
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('06583')
```

### Seed 08366
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('08366')
```

## Populations that crossed once

### Seed 00833
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('00833')
```

### Seed 01357
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('01357')
```

### Seed 02290
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('02290')
```

### Seed 02359
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('02359')
```

### Seed 03149
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('03149')
```

### Seed 05295
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('05295')
```

### Seed 07051
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('07051')
```

### Seed 07605
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('07605')
```

### Seed 07916
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('07916')
```

### Seed 09839
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('09839')
```


## Populations that crossed twice

Note that this is the seed in Figure 7 of the paper.

### Seed 05501
```{r, message=FALSE, warning=FALSE}
plot_replay_with_adjusted_benchmark('05501')
```